NN_Accel_Heading

### 2024-03-07 Fish Data Analysis/NN_Accel_Heading.qmd
Cleaned up rad and degrees, checked functions

@@

fish_comp_data_switch_fish <- fish_comp_data %>% mutate(Fish3 = Fish1,
                                              Fish3_X = Fish1_X,
                                              Fish3_Y = Fish1_Y,
                                              Fish3_Z = Fish1_Z,
                                              Fish3_Speed = Fish1_Speed,
                                              Fish3_Accel = Fish1_Accel,
                                              Fish3_Yaw_Heading = Fish1_Yaw_Heading,
                                              Fish3_Yaw_Heading_PS = Fish1_Yaw_Heading_PS
                                              ) %>%
                                       mutate(Fish1 = Fish2,
                                              Fish1_X = Fish2_X,
                                              Fish1_Y = Fish2_Y,
                                              Fish1_Z = Fish2_Z,
                                              Fish1_Speed = Fish2_Speed,
                                              Fish1_Accel = Fish2_Accel,
                                              Fish1_Yaw_Heading = Fish2_Yaw_Heading,
                                              Fish1_Yaw_Heading_PS = Fish2_Yaw_Heading_PS) %>%
                                       mutate(Fish2 = Fish3,
                                              Fish2_X = Fish3_X,
                                              Fish2_Y = Fish3_Y,
                                              Fish2_Z = Fish3_Z,
                                              Fish2_Speed = Fish3_Speed,
                                              Fish2_Accel = Fish3_Accel,
                                              Fish2_Yaw_Heading = Fish3_Yaw_Heading,
                                              Fish2_Yaw_Heading_PS = Fish3_Yaw_Heading_PS) %>%
                                       #Switch around angle when we switch focal fish
                                       mutate(Angle = ifelse(Angle < 0, Angle + 180, Angle - 180),
                                              Relative_X = -1*Relative_X) %>%
                                       select(-c(Fish3,Fish3_X,Fish3_Y,Fish3_Z,Fish3_Speed,Fish3_Accel,Fish3_Yaw_Heading,Fish3_Yaw_Heading_PS))

@@

fish_data_nn <- fish_comp_data_combo %>% group_by(Fish1,Tailbeat_Num,Data_ID) %>%
                                    mutate(nndDistance = min(Distance, na.rm = TRUE),
                                           nndDistance = ifelse(is.infinite(nndDistance), NA, nndDistance),
                                           all_nas = all(is.na(nndDistance))) %>%
                                    filter(Distance == nndDistance | all_nas) %>%
                                    mutate(row_n = row_number()) %>%
                                    filter(row_n == 1) %>%
                                    ungroup() %>%
                                    mutate(Angle_Rads = deg2rad(Angle),
                                           Fish1_Yaw_Heading_PS = deg2rad(Fish1_Yaw_Heading_PS)) %>%
                                    filter(Distance <= 2)

nn_heading_sum <- fish_data_nn %>% mutate(angle_bins = round_any(Angle_Rads,pi/8)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

nn_accel_sum <- fish_data_nn %>% mutate(distance_bins = round_any(Relative_X,0.25)) %>%
                                                group_by(distance_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_accel = mean(Fish1_Accel, na.rm = T),
                                                          sd_accel = sd(Fish1_Accel, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

@@

heading_ps_angle_step <- stepAIC(heading_ps_angle_base,
                                   list(upper = ~ ((cos(Angle_Rads)+sin(Angle_Rads))*Ablation*Darkness), lower = ~1),
                                   direction = "both",
                                   trace = F)

@@

tester_model <- lm(Fish1_Yaw_Heading_PS ~ sin(Angle_Rads)*Ablation*Darkness, data = fish_data_nn_model) 
Anova(tester_model)

ggplot(fish_data_nn %>% filter(Flow == "Still Water"),
       aes(x = Angle_Rads, y = Fish1_Yaw_Heading_PS, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  #geom_smooth(se = F)+
  geom_smooth(se = F, method = glm, formula = y ~ 1)+
  #geom_line(aes(y = predict(heading_ps_angle_step)), size = 1)+
  theme_classic() +
  #ylim(-5,5)+
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")

ggplot(nn_heading_sum %>% filter(Flow == "Still Water"),
       aes(x = angle_bins, y = rad2deg(mean_turn), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  #geom_smooth(se = F)+
  #geom_line(aes(y = sin_pred)) +
  #geom_line(alpha = 0.75) +
  #geom_smooth(se = F, method = glm, formula = y ~ sin(x))+
  geom_line(aes(y = rad2deg(predict(heading_ps_angle_step,
                    newdata = nn_heading_sum %>% filter(Flow == "Still Water") %>% mutate(Angle_Rads = angle_bins)))), size = 1)+
  theme_classic() +
  #ylim(-0.6,0.6)+
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (degrees/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  # scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
  #                    labels = c("-π","-3π/4","-π/2","-π/4","0","π/4","π/2","3π/4","π"))+
  labs(color = "Sensory Conditions")

>>>>

```{r}
raw_new <- read.csv("Data/Fish_Comp_Values_3D.csv")
raw_old <- read.csv("Data/older_Fish_Comp_Values_3D.csv") %>% mutate(Angle = deg2rad(Angle),
                                                                     Fish1_Yaw_Heading = deg2rad(Fish1_Yaw_Heading),
                                                                     Fish1_Yaw_Heading_PS = deg2rad(Fish1_Yaw_Heading_PS))

hist(raw_new$Angle - deg2rad(raw_old$Angle))
hist(raw_new$Fish1_Yaw_Heading - deg2rad(raw_old$Fish1_Yaw_Heading))
hist(raw_new$Fish1_Yaw_Heading_PS - deg2rad(raw_old$Fish1_Yaw_Heading_PS))

@@

fish_comp_data_switch_fish <- fish_comp_data %>% mutate(Fish3 = Fish1,
                                              Fish3_X = Fish1_X,
                                              Fish3_Y = Fish1_Y,
                                              Fish3_Z = Fish1_Z,
                                              Fish3_Speed = Fish1_Speed,
                                              Fish3_Accel = Fish1_Accel,
                                              Fish3_Yaw_Heading = Fish1_Yaw_Heading,
                                              Fish3_Yaw_Heading_PS = Fish1_Yaw_Heading_PS
                                              ) %>%
                                       mutate(Fish1 = Fish2,
                                              Fish1_X = Fish2_X,
                                              Fish1_Y = Fish2_Y,
                                              Fish1_Z = Fish2_Z,
                                              Fish1_Speed = Fish2_Speed,
                                              Fish1_Accel = Fish2_Accel,
                                              Fish1_Yaw_Heading = Fish2_Yaw_Heading,
                                              Fish1_Yaw_Heading_PS = Fish2_Yaw_Heading_PS) %>%
                                       mutate(Fish2 = Fish3,
                                              Fish2_X = Fish3_X,
                                              Fish2_Y = Fish3_Y,
                                              Fish2_Z = Fish3_Z,
                                              Fish2_Speed = Fish3_Speed,
                                              Fish2_Accel = Fish3_Accel,
                                              Fish2_Yaw_Heading = Fish3_Yaw_Heading,
                                              Fish2_Yaw_Heading_PS = Fish3_Yaw_Heading_PS) %>%
                                       #Switch around angle when we switch focal fish
                                       mutate(Angle = ifelse(Angle >= 0, Angle - pi, Angle + pi),
                                              Relative_X = -1*Relative_X) %>%
                                       select(-c(Fish3,Fish3_X,Fish3_Y,Fish3_Z,Fish3_Speed,Fish3_Accel,Fish3_Yaw_Heading,Fish3_Yaw_Heading_PS))

@@

fish_data_nn <- fish_comp_data_combo %>% group_by(Fish1,Tailbeat_Num,Data_ID) %>%
                                    mutate(nndDistance = min(Distance, na.rm = TRUE),
                                           nndDistance = ifelse(is.infinite(nndDistance), NA, nndDistance),
                                           all_nas = all(is.na(nndDistance))) %>%
                                    filter(Distance == nndDistance | all_nas) %>%
                                    mutate(row_n = row_number()) %>%
                                    filter(row_n == 1) %>%
                                    ungroup() %>%
                                    mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS) %>%
                                    filter(Distance <= 2)

fish_data_nn_normal <- fish_data_nn %>% group_by(Flow_Ablation_Darkness) %>%
                                        mutate(Fish1_Yaw_Heading_PS = Fish1_Yaw_Heading_PS - mean(Fish1_Yaw_Heading_PS)) %>%
                                        ungroup()

nn_heading_sum <- fish_data_nn_normal %>% mutate(angle_bins = round_any(Angle,pi/6)) %>%
                                                group_by(angle_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_turn = mean(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          sd_turn = sd(Fish1_Yaw_Heading_PS, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

nn_accel_sum <- fish_data_nn_normal %>% mutate(distance_bins = round_any(Relative_X,0.25)) %>%
                                                group_by(distance_bins,Flow,Ablation,Darkness,Flow_Ablation_Darkness) %>%
                                                summarise(mean_accel = mean(Fish1_Accel, na.rm = T),
                                                          sd_accel = sd(Fish1_Accel, na.rm = T),
                                                          n = n()) %>%
                                                ungroup() %>%
                                                na.omit()

@@

heading_ps_angle_step <- stepAIC(heading_ps_angle_base,
                                   list(upper = ~ (sin(Angle)+Ablation:sin(Angle)+Darkness:sin(Angle)), lower = ~1),
                                   direction = "both",
                                   trace = F)

@@

tester_model <- lm(Fish1_Yaw_Heading_PS ~ sin(Angle):Ablation:Darkness, data = fish_data_nn_model) 
Anova(tester_model)

ggplot(fish_data_nn %>% filter(Flow == "Still Water"),
       aes(x = Angle, y = Fish1_Yaw_Heading_PS, color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point(alpha = 0.1)+
  facet_wrap(~ interaction(Ablation, Darkness, sep =", ")) +
  #geom_smooth(se = F)+
  #geom_smooth(se = F, method = glm, formula = y ~ 1)+
  geom_line(aes(y = predict(heading_ps_angle_step)), size = 1)+
  theme_classic() +
  #ylim(-5,5)+
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (rad/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  labs(color = "Sensory Conditions")

ggplot(nn_heading_sum %>% filter(Flow == "Still Water"),
       aes(x = angle_bins, y = rad2deg(mean_turn), color = interaction(Ablation, Darkness, sep =", ")))+
  geom_point()+
  #geom_smooth(se = F)+
  #geom_line(aes(y = sin_pred)) +
  #geom_line(alpha = 0.75) +
  #geom_smooth(se = F, method = glm, formula = y ~ sin(x))+
  geom_line(aes(y = rad2deg(predict(tester_model,
                    newdata = nn_heading_sum %>% filter(Flow == "Still Water") %>% mutate(Angle = angle_bins)))), size = 1)+
  theme_classic() +
  ylim(-50,50) +
  xlab("direction of neighbour (degrees)") +
  ylab("focal fish turning angle (degrees/sec)") +
  scale_colour_manual(values = c(light_no_ab_color,light_ab_color,dark_no_ab_color,dark_ab_color)) +
  scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
                     labels = rad2deg(unique(nn_heading_sum$angle_bins))) +
  # scale_x_continuous(breaks = unique(nn_heading_sum$angle_bins),
  #                    labels = c("-π","-3π/4","-π/2","-π/4","0","π/4","π/2","3π/4","π"))+
  labs(color = "Sensory Conditions")

# emmeans(tester_model, ~ Ablation|Darkness, type = "response")
# emmeans(tester_model, ~ Darkness|Ablation, type = "response")
